TODO LIST

### 2019-03-01 LMCS/sendrev/response-letter.tex

	A proof was added.

%	\item[$\checkmark$] p.31, definition of $\leq$ on $\F$: ``$f(i) \leq g(i)$ for every $i \in \{1,\ldots,\ell\}$'': Inputs for $f$ and $g$ should be elements of $A^{\ell}$, not $\{1,\ldots,\ell\}$
	\item \red{p.32/33, example 6.2: I don't see why the formula given for dag's does not work for arbitrary graphs: In arbitrary graphs, the fixpoint is still reached after at most $n-1$ steps, since the maximal length of any shortest path in any graph is $n-1$. On the other hand, assuming that I am mistaken and the support does not become stable within $n-1$ steps: Using the formula	given for arbitrary graphs, the support of the constructed function restricted to those tuples that have first in their last component is the same as the support of the constructed function using the formula for dags in each step of the recursion. Hence, the formula should not be able to fix such a problem.}

	The problem with the first formula for directed graphs is precisely the opposite. The fixed point could be reached too soon without counting all paths of length at most $n-1$. The fixed point is always reached after $k$ steps, where $k$ is the size of the maximum shortest path between two vertices in the graph. If the graph has a cycle in it, there might be paths of size $k' \in \{k+1,\ldots,n-1\}$ that would not be counted because the fixed point was reached earlier. That is why we need a different formula for the general case. We added a comment about this in the paper.

### 2019-03-01 LMCS/sendrev/beyond.tex
sc

\begin{exa} \label{ex:count-path}
We would like to define an $\rqfo(\fo)$-formula that, given a directed acyclic graph $G$ with $n$ nodes and a pair of nodes $b$, $c$ in $G$, counts the number of paths of length less than $n$ from $b$ to $c$ in $G$. To this end, assume that graphs are encoded using the signature $\R = \{ E(\cdot,\cdot)\}$, and then define formula $\alpha(x, y, f)$ as follows:
$$ 
E(x,y) + \sa{z} f(x,z)\cdot E(z,y).
$$
We have that $\clfp{\alpha(x,y,f)}$ defines our counting function. In fact, assume that $\A$ is an $\R$-structure with $n$ elements in its domain encoding an acyclic directed graph. Moreover, assume that $b,c$ are elements of $\A$ and $v$ is a first-order assignment over $\A$ such that $v(x) = b$ and $v(y) = c$. Then we have that $\sem{\clfp{\alpha(x,y,f)}}(\A,v)$ is equal to the number of paths in $\A$ from $b$ to $c$ of length at most $n-1$.

Assume now that we need to extend our previous counting function to the case of arbitrary directed graphs. To this end, suppose that $\varphi_{\text{\rm first}}(x)$ and $\varphi_{\text{succ}}(x,y)$ are $\fo$-formulae defining the first element of $<$ and the successor relation associated to $<$, respectively.\footnote{Recall that in this paper we consider ordered $\R$-structures.} Moreover, define formula $\beta(x, y, t, g)$ as follows:
\begin{align*}
(E(x,y) + \sa{z} g(x,z,t)\cdot E(z,y)) \cdot \varphi_{\text{\rm first}}(t) \ +
\sa{t'} \varphi_{\text{succ}}(t',t) \cdot \left(\sa{x'} \sa{y'} g(x',y',t') \right)
\end{align*}
Then our extended counting function is defined by:
$$
\sa{t} (\varphi_{\text{\rm first}}(t) \cdot \clfp{\beta(x,y,t,g)}).
$$ 
In fact, the number of paths of length at most $n$ from a node $x$ to a node $y$ is recursively computed by using the formula $(E(x,y) + \sa{z} g(x,z,t)\cdot E(z,y)) \cdot \varphi_{\text{\rm first}}(t)$, which stores this value in $g(x,y,t)$ with $t$ the first element in the domain.  The other formula $\sa{t'} \varphi_{\text{succ}}(t',t) \cdot \left(\sa{x'} \sa{y'} g(x',y',t') \right)$ is just an auxiliary artifact that is used as a counter to allow reaching a fixed point in the support of $g$ in $n$ steps. Notice that the use of the filter $\varphi_{\text{succ}}(t',t)$ prevents this formula for incrementing the value of $g(x,y,t)$ when $t$ is the first element in the domain.
\end{exa}
\martin{El Reviewer 1 hizo un comentario de este ejemplo, lo respond\'i en la carta}
In contrast to $\lfp$, to reach a fixed point we do not need to impose any positive restriction on the formula $\beta(\x,h)$.
In fact, since $\beta$ is constructed from monotone operations (sum and product) over the natural numbers, the resulting operator $T_{\beta}$ is monotone as well.

>>>>

\begin{exa} \label{ex:count-path}
We would like to define an $\rqfo(\fo)$-formula that, given a directed acyclic graph $G$ with $n$ nodes and a pair of nodes $b$, $c$ in $G$, counts the number of paths of length less than $n$ from $b$ to $c$ in $G$. To this end, assume that graphs are encoded using the signature $\R = \{ E(\cdot,\cdot),<\}$, and then define formula $\alpha(x, y, f)$ as follows:
\begin{eqnarray}\label{eq-count-paths-acyclic}
E(x,y) + \sa{z} f(x,z)\cdot E(z,y).
\end{eqnarray}
We have that $\clfp{\alpha(x,y,f)}$ defines our counting function. In fact, assume that $\A$ is an $\R$-structure with $n$ elements in its domain encoding an acyclic directed graph. Moreover, assume that $b,c$ are elements of $\A$ and $v$ is a first-order assignment over $\A$ such that $v(x) = b$ and $v(y) = c$. Then we have that $\sem{\clfp{\alpha(x,y,f)}}(\A,v)$ is equal to the number of paths in $\A$ from $b$ to $c$ of length at most $n-1$.

Assume now that we need to count the number of paths of length less than $n$ from a node $b$ to a node $c$ in a directed graph that is not necessarily acyclic. 
%extend our previous counting function to the case of arbitrary directed graphs. 
Then we cannot use formula \eqref{eq-count-paths-acyclic}, as the fixed point could be reached too soon without counting all paths of length at most $n-1$. The fixed point of \eqref{eq-count-paths-acyclic} is always reached after $k$ steps, where $k$ is the size of the maximum shortest path between two vertices in the graph. If the graph has a cycle in it, there might be paths of size $k' \in \{k+1,\ldots,n-1\}$ that would not be counted because the fixed point was reached earlier. Thus, we need a more involved formula in the general case, which is given below.

%To this end, 
Suppose that $\varphi_{\text{\rm first}}(x)$ and $\varphi_{\text{succ}}(x,y)$ are $\fo$-formulae defining the first element of $<$ and the successor relation associated to $<$, respectively.
%\footnote{Recall that in this paper we consider ordered $\R$-structures.} 
Moreover, define formula $\beta(x, y, t, g)$ as follows:
\begin{align*}
(E(x,y) + \sa{z} g(x,z,t)\cdot E(z,y)) \cdot \varphi_{\text{\rm first}}(t) \ +
\sa{t'} \varphi_{\text{succ}}(t',t) \cdot \left(\sa{x'} \sa{y'} g(x',y',t') \right)
\end{align*}
Then our extended counting function is defined by:
$$
\sa{t} (\varphi_{\text{\rm first}}(t) \cdot \clfp{\beta(x,y,t,g)}).
$$ 
In fact, the number of paths of length at most $n$ from a node $x$ to a node $y$ is recursively computed by using the formula $(E(x,y) + \sa{z} g(x,z,t)\cdot E(z,y)) \cdot \varphi_{\text{\rm first}}(t)$, which stores this value in $g(x,y,t)$ with $t$ the first element in the domain.  The other formula $\sa{t'} \varphi_{\text{succ}}(t',t) \cdot \left(\sa{x'} \sa{y'} g(x',y',t') \right)$ is just an auxiliary artifact that is used as a counter to allow reaching a fixed point in the support of $g$ in $n$ steps. Notice that the use of the filter $\varphi_{\text{succ}}(t',t)$ prevents this formula for incrementing the value of $g(x,y,t)$ when $t$ is the first element in the domain.
\qed
\end{exa}
\martin{El Reviewer 1 hizo un comentario de este ejemplo, lo respond\'i en la carta}
\marcelo{Me parece bien la respuesta en la carta. Agregue esta respuesta en el paper, ayuda a entender mejor el ejemplo.}
In contrast to $\lfp$, to reach a fixed point we do not need to impose any positive restriction on the formula $\beta(\x,h)$.
In fact, since $\beta$ is constructed from monotone operations (sum and product) over the natural numbers, the resulting operator $T_{\beta}$ is monotone as well.

### 2019-03-01 LMCS/sendrev/todo-list.tex
sc

\item[?] p.32/33, example 6.2: I don't see why the formula given for dag's does not work for arbitrary

@@

>>>>

\item[$\checkmark$] p.32/33, example 6.2: I don't see why the formula given for dag's does not work for arbitrary

@@

\marcelo{Esta bien la respuesta en la carta. La agregue al ejemplo para que quedara mas claro}
